Digit Recognition: Nearest Neighbors vs. Perceptron

### 2013-06-10 Paper/paper.tex
First draft of the paper done

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{hyperref}
\usepackage[margin=1.5in]{geometry}
%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

@@

\begin{table}[ht]
\textbf{\caption{Subset of Kaggle Data}} % title
\centering 
\begin{tabular}{c | c c c }
Label & Pixel 77 & Pixel 78 & Pixel 79 \\
\hline
8 & 0 & 0 & 0 \\
6 & 89 & 208 & 135 \\ [1ex]
\end{tabular}
\label{table:nonlin}
\end{table}

@@

@@

\subsection{Runtime}

@@

\section{Kernelized Perceptron}
\subsubsection{Overview}

\subsubsection{Implementation Details}

\subsubsection{Results}

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone, regardless of the formatter being used.

\subsection{Citations within the text}

Citations within the text should be numbered consecutively. The corresponding
number is to appear enclosed in square brackets, such as [1] or [2]-[5]. The
corresponding references are to be listed in the same order at the end of the
paper, in the \textbf{References} section. (Note: the standard
\textsc{Bib\TeX} style \texttt{unsrt} produces this.) As to the format of the
references themselves, any style is acceptable as long as it is used
consistently.

As submission is double blind, refer to your own published work in the 
third person. That is, use ``In the previous work of Jones et al.\ [4]'',
not ``In our previous work [4]''. If you cite your other papers that
are not widely available (e.g.\ a journal paper under review), use
anonymous author names in the citation, e.g.\ an author of the
form ``A.\ Anonymous''.

\subsection{Footnotes}

Indicate footnotes with a number\footnote{Sample of the first footnote} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches
(12~picas).\footnote{Sample of the second footnote}

\subsection{Figures}

All artwork must be neat, clean, and legible. Lines should be dark
enough for purposes of reproduction; art work should not be
hand-drawn. The figure number and caption always appear after the
figure. Place one line space before the figure caption, and one line
space after the figure. The figure caption is lower case (except for
first word and proper nouns); figures are numbered consecutively.

Make sure the figure caption does not get separated from the figure.
Leave sufficient space to avoid splitting the figure and figure caption.

You may use color figures. 
However, it is best for the
figure captions and the paper body to make sense if the paper is printed
either in black/white or in color.
\begin{figure}[h]
\begin{center}
%\framebox[4.0in]{$\;$}
\fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
\end{center}
\caption{Sample figure caption.}
\end{figure}

\subsection{Tables}

All tables must be centered, neat, clean and legible. Do not use hand-drawn
tables. The table number and title always appear before the table. See
Table~\ref{sample-table}.

Place one line space before the table title, one line space after the table
title, and one line space after the table. The table title must be lower case
(except for first word and proper nouns); tables are numbered consecutively.

\begin{table}[t]
\caption{Sample table title}
\label{sample-table}
\begin{center}
\begin{tabular}{ll}
\multicolumn{1}{c}{\bf PART}  &\multicolumn{1}{c}{\bf DESCRIPTION}
\\ \hline \\
Dendrite         &Input terminal \\
Axon             &Output terminal \\
Soma             &Cell body (contains cell nucleus) \\
\end{tabular}
\end{center}
\end{table}

\section{Final instructions}
Do not change any aspects of the formatting parameters in the style files.
In particular, do not modify the width or length of the rectangle the text
should fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PostScript or PDF files}

Please prepare PostScript or PDF files with paper size ``US Letter'', and
not, for example, ``A4''. The -t
letter option on dvips will produce US Letter files.

Fonts were the main cause of problems in the past years. Your PDF file must
only contain Type 1 or Embedded TrueType fonts. Here are a few instructions
to achieve this.

\begin{itemize}

\item You can check which fonts a PDF files uses.  In Acrobat Reader,
select the menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts
are also acceptable for NIPS. Please see
\url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item LaTeX users:

\begin{itemize}

\item Consider directly generating PDF files using \verb+pdflatex+
(especially if you are a MiKTeX user). 
PDF figures must be substituted for EPS figures, however.

\item Otherwise, please generate your PostScript and PDF files with the following commands:
\begin{verbatim} 
dvips mypaper.dvi -t letter -Ppdf -G0 -o mypaper.ps
ps2pdf mypaper.ps mypaper.pdf
\end{verbatim}

Check that the PDF files only contains Type 1 fonts. 
%For the final version, please send us both the Postscript file and
%the PDF file.

\item xfig "patterned" shapes are implemented with 
bitmap fonts.  Use "solid" shapes instead. 
\item The \verb+\bbold+ package almost always uses bitmap
fonts.  You can try the equivalent AMS Fonts with command
\begin{verbatim}
\usepackage[psamsfonts]{amssymb}
\end{verbatim}
 or use the following workaround for reals, natural and complex: 
\begin{verbatim}
\newcommand{\RR}{I\!\!R} %real numbers
\newcommand{\Nat}{I\!\!N} %natural numbers 
\newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}

\item Sometimes the problematic fonts are used in figures
included in LaTeX files. The ghostscript program \verb+eps2eps+ is the simplest
way to clean such figures. For black and white figures, slightly better
results can be achieved with program \verb+potrace+.
\end{itemize}
\item MSWord and Windows users (via PDF file):
\begin{itemize}
\item Install the Microsoft Save as PDF Office 2007 Add-in from
\url{http://www.microsoft.com/downloads/details.aspx?displaylang=en\&familyid=4d951911-3e7e-4ae6-b059-a2e79ed87041}
\item Select ``Save or Publish to PDF'' from the Office or File menu
\end{itemize}
\item MSWord and Mac OS X users (via PDF file):
\begin{itemize}
\item From the print menu, click the PDF drop-down box, and select ``Save
as PDF...''
\end{itemize}
\item MSWord and Windows users (via PS file):
\begin{itemize}
\item To create a new printer
on your computer, install the AdobePS printer driver and the Adobe Distiller PPD file from
\url{http://www.adobe.com/support/downloads/detail.jsp?ftpID=204} {\it Note:} You must reboot your PC after installing the
AdobePS driver for it to take effect.
\item To produce the ps file, select ``Print'' from the MS app, choose
the installed AdobePS printer, click on ``Properties'', click on ``Advanced.''
\item Set ``TrueType Font'' to be ``Download as Softfont''
\item Open the ``PostScript Options'' folder
\item Select ``PostScript Output Option'' to be ``Optimize for Portability''
\item Select ``TrueType Font Download Option'' to be ``Outline''
\item Select ``Send PostScript Error Handler'' to be ``No''
\item Click ``OK'' three times, print your file.
\item Now, use Adobe Acrobat Distiller or ps2pdf to create a PDF file from
the PS file. In Acrobat, check the option ``Embed all fonts'' if
applicable.
\end{itemize}

\end{itemize}
If your file contains Type 3 fonts or non embedded TrueType fonts, we will
ask you to fix it.

\subsection{Margins in LaTeX}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+
from the graphicx package. Always specify the figure width as a multiple of
the line width as in the example below using .eps graphics
\begin{verbatim}
   \usepackage[dvips]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.eps} 
\end{verbatim}
or % Apr 2009 addition
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ... 
   \includegraphics[width=0.8\linewidth]{myfile.pdf} 
\end{verbatim}
for .pdf graphics. 
See section 4.4 in the graphics bundle documentation (\url{http://www.ctan.org/tex-archive/macros/latex/required/graphics/grfguide.ps})

A number of width problems arise when LaTeX cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All
acknowledgments go at the end of the paper. Do not include 
acknowledgments in the anonymized submission, only in the 
final paper. 
\subsubsection*{References}

References follow the acknowledgments. Use unnumbered third level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to `small' (9-point) 
when listing the references. {\bf Remember that this year you can use
a ninth page as long as it contains \emph{only} cited references.}

\small{
[1] Alexander, J.A. \& Mozer, M.C. (1995) Template-based algorithms
for connectionist rule extraction. In G. Tesauro, D. S. Touretzky
and T.K. Leen (eds.), {\it Advances in Neural Information Processing
Systems 7}, pp. 609-616. Cambridge, MA: MIT Press.

[2] Bower, J.M. \& Beeman, D. (1995) {\it The Book of GENESIS: Exploring
Realistic Neural Models with the GEneral NEural SImulation System.}
New York: TELOS/Springer-Verlag.

[3] Hasselmo, M.E., Schnell, E. \& Barkai, E. (1995) Dynamics of learning
and recall at excitatory recurrent synapses and cholinergic modulation
in rat hippocampal region CA3. {\it Journal of Neuroscience}
{\bf 15}(7):5249-5262.
}

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips12submit_e,times}
\usepackage{hyperref}
\usepackage[margin=1.5in]{geometry}
\usepackage{graphicx}

%\documentstyle[nips12submit_09,times,art10]{article} % For LaTeX 2.09

@@

\begin{table}[ht]
\textbf{\caption{Subset of Kaggle Data}} % title
\centering 
\begin{tabular}{c | c c c }
Label & Pixel 77 & Pixel 78 & Pixel 79 \\
\hline
8 & 0 & 0 & 0 \\
6 & 89 & 208 & 135 \\ [1ex]
\end{tabular}
\label{table:Data}
\end{table}

@@

\begin{center}
\includegraphics[width=406px]{KNN_CrossValidation.png}
\end{center}

@@

@@

Our KNN implementation was extremely successful, correctly classifying 96.857\%
of the testing samples. One downside to it however is the runtime. Training is
near instantaneous since the algorithm simply loads the data into an array. To
classify a sample the algorithm needs to compute the euclidean distance between
the testing sample and every training sample. In the case of the full data set
this is 42,000 samples, each of which has 768 dimensions. As a result, each
testing sample takes approximately. Altogether it takes close to 8 hours to
classify all 28,000 testing samples.

\section{Kernelized Perceptron}
\subsection{Overview}

The perceptron algorithm attempts to separate a training data set into two
classes and attempts to classify future testing samples as belonging to one of
these two classes. It does this by storing a list of all the mistakes it has
made previously and using this knowledge to adjust future predictions in the
hopes of improving classification accuracy.

In its most basic implementation, the perceptron algorithm divides the data
linearly. However, the data is often not linearly separable. To address this
issue we use a kernel function to create a non-linear decision boundary. This
allows us to have a more complex decision boundary which potentially fits the
data better.

Before we can begin classifying test points, we need to train the perceptron
algorithm using our training data set. The training process attempts
to classify each training point sequentially. If the perceptron correctly
classifies the point nothing changes and the training process moves on to the
next training sample. If the training sample is incorrectly classified, however,
the mistake is added to an array of mistakes which is taken into consideration
for future classifications.

One problem that needs to be addressed is that a standard perceptron is a binary
classifier meaning it selects between one of two classes. We resolved this issue
by training $10 \choose 2 = 45$ classifiers. Each classifier is responsible for
classifying between two digits. For example, we would train a perceptron to
classify between 0 and 1, and a second to classify between 0 and 2, and so
on. When we classify we vote between all of our trained classifiers and output
the majority element.

The actual classification is performed by:

\begin{center}
$classify(X) = \sum\limits_{i \in Mistakes}{Y_iK(X_i, X)}$
\end{center}

The kernel function is represented by $K(X_i, X)$. There are a number of choices
for this kernel function.

\subsection{Cross Validation for Kernel Selection}

We implemented two kernel functions to use with our perceptron algorithm. One is
a polynomial kernel which takes the form:

\begin{center}
$Polynomial(x, y, d) = (dot(x, y) + 1)^d$
\end{center}

The second is an exponential kernel:

\begin{center}
$Exponential(x, y, \sigma) = exp(-\frac{||x - y||}{2\sigma^2})$
\end{center}

Both of these kernels have a parameter which needs to be tuned, $d$ in the case of
the polynomial kernel and $\sigma$ in the case of the exponential kernel. We
used cross validation to find out which kernel gave us the best result as well
as which value of $d$ and $\sigma$ was optimal.

\begin{table}[ht]
\textbf{\caption{Perceptron Cross Validation Success Rates}} % title
\centering
\begin{tabular}{c | c c }
Kernel & $d, \sigma = 5$ & $d, \sigma = 10$ \\
\hline
Polynomial & 95.67\% & 94.52\% \\
Exponential & 93.84\% & 96.3\% \\ [1ex]
\end{tabular}
\label{table:perceptron}
\end{table}

\begin{center}
\includegraphics[width=406px]{Perceptron_CrossValidation.png}
\end{center}

Our cross validation indicates that the exponential kernel with a $\sigma$ of 10
leads to the highest success rate of 96.3\%. An interesting observation is that
our polynomial kernel actually became less successful when it's dimension was
increased whereas the exponential kernel got better. We suspect that this is
because the polynomial kernel began to overfit the training data.

Finally, the runtime of the perceptron algorithm is a little over an hour. It
takes approximately 45 minutes to train and about 20 minutes to test the
data. The faster runtime allowed us to cross validate on the full data set.

\section{Conclusion}
Despite its seeming simplicity, we found that KNN was extremely successful
when classifying data points in the Kaggle set.

\begin{center}
\begin{tabular}{ c| c c}
	\hline
Algorithm & Percent Success & runtime\\
	\hline
Nearest Neighbors, $k = 4$   & 96.857\% & $\sim$8 hours\\
Perceptron   & 96.300\%  & $\sim$1 hour\\
\end{tabular}
\end{center}

Table 4 summarizes the accuracy and runtimes of weighted KNN with $K = 4$ and
perceptron with the polynomial kernel and $d = 10$. KNN is clearly more
accurate, however because perceptron is 8x faster and almost as accurate, there
could be many cases where it is a better choice.

In addition to being faster, perceptron likely has a lower memory
footprint. While KNN must store every sample in the training set, perceptron
only stores mistakes which will be a subset of the entire training set.

\section{Future Work}
There are three specific expansions to our work we would have liked to explore
had we had time.

\begin{itemize}
\item Expanded cross validation for KNN using the full training set. This would
  help validate that the weighted voting algorithm with $K = 4$ leads to the
  most accurate. 
\item More kernels and a greater assortment of $\sigma$ and $d$ values to
  see if we can increase the effectiveness of the perceptron algorithm. 
\item An SVM implementation which regularizes the data to avoid overfitting. If
  the perceptron implementation overfits the data, this would be a potential
  way to increase the accuracy.
\end{itemize}

\begingroup
\renewcommand{\section}[2]{}%
%\renewcommand{\chapter}[2]{}% for other classes
\bibliographystyle{unsrt}
\bibliography{references}	
\endgroup
